---
title: "Cleaning Data with R"
output: html_notebook
---

> First the library and data need to be loaded into R Note: This walkthrough is from
https://towardsdatascience.com/data-cleaning-with-r-and-the-tidyverse-detecting-missing-values-ea23c519bc62 

```{r}
library(tidyverse)

#set the working directory
file_loc <- paste(getwd(), "/raw_data", sep="")
setwd(file_loc)

# read csv file
df <- read_csv("telecom.csv")
```

> The "glimpse" function shows the variable names and types.

```{r}
glimpse(df)
```

> The "dplyr" library is used for data manipulation. The first example will demonstrate the use of filter to show all customers who have Churn values equal to "yes".

```{r}
df %>%
  filter(Churn == "yes")
```

> Filter and select can be used together to filter out those who have churned and only select the customerID and the TotalCharges. This is also completed by piping the commands and the dplyr package.

```{r}
# filter on customers that churned,
# select customerID and TotalCharges columns
df %>%
  filter(Churn=="yes") %>%
  select(customerID, TotalCharges)
```

> Now the missing values can be tackled. I'll look at the monthly charges column by using the same select method.

```{r}
# select MonthlyCharges Column
df %>%
  select(MonthlyCharges)
```

> There are three missing values. One of them is NaN and the others a blank. Using is.na, lets find outh ow R recognizes these values.

```{r}
# looking at MonthlyCharges
df$MonthlyCharges
is.na(df$MonthlyCharges)
```

> Judging by the results all missing values are recognized as missing values

> Using distinct, we can see how many and which distinct values there are in the column

```{r}
# looking at the distinct values
df %>%
  distinct(MonthlyCharges)
```

> NA only shows up once because there are two values and we wanted all unique values

> To get a count we can use summarise.

```{r}
# counting unique values
df %>%
  summarise(n = n_distinct(MonthlyCharges))
```

> This returns a tibble with a column that we named "n" for the count of distinct values

> A similar function may be used to count the total number of NA values using summarise and is.na

```{r}
# counting missing values
df %>%
  summarise(count = sum(is.na(MonthlyCharges)))
```

> Both of the previous actions can be grouped up into one function and we can also calculate the median charges, as well.

```{r}
# counting unique, missing, and median values
df %>% summarise(n = n_distinct(MonthlyCharges),
                 na = sum(is.na(MonthlyCharges)),
                 med = median(MonthlyCharges, na.rm = TRUE))
```

> Now that the missing values have been identified they can be replaced with the median value of MonthlyCharges by using the mutate function.

```{r}
# mutate missing values
df %>%
  mutate(MonthlyCharges
         = replace(MonthlyCharges,
                   is.na(MonthlyCharges),
                   median(MonthlyCharges, na.rm = TRUE)))
```

> Double check that this was applied to the entire tibble.

```{r}
df
```

> It didn't change. That's because the mutation was not assigned to the original dataframe. The below code will fix this.

```{r}
# mutate missing values and modify the original dataframe
df <- df %>%
  mutate(MonthlyCharges = replace(MonthlyCharges,
                                  is.na(MonthlyCharges),
                                  median(MonthlyCharges, na.rm=TRUE)))
```

> Now double check again to make sure this process worked.

```{r}
df
```

> It worked!

> Now it's time to tackle some non-standard missing values

> To start, take a look at the TotalCharges column

```{r}
# look at the TOtalCharges column
df %>%
  select(TotalCharges)
```

> At a glance there are various forms of the NA value. The next step is to see if R can properly interpret these values as NA using the is.na function.

```{r}
# looking at missing values
df$TotalCharges
is.na(df$TotalCharges)
```

> Ok, so looking at the results there are three na values and one TRUE value. I looks lke the only value of NA that R recognizes is NA.

> Next, we summarise the data to see how many missing values R found.

```{r}
# counting missing values
df %>%
  summarise(count = sum(is.na(TotalCharges)))
```

> As expected from the previous test. R recognises only one missing value.

> The next step is to use the mutate function to change the other missing values to the correct format.

```{r}
# replacig with standard missing value type, NA
df <- df %>%
  mutate(TotalCharges = replace(TotalCharges, TotalCharges == "na", NA)) %>%
  mutate(TotalCharges = replace(TotalCharges, TotalCharges == "N/A", NA))
```

> Check the colum again to varify the changes took place.

```{r}
# take another look
df$TotalCharges
is.na(df$TotalCharges)
```

> Great! The changes were made successfully. There is still an issue with this column. R is interpreting the values as characters rather than a numerical value. To verify use the glimpse function.

```{r}
# verify column type
glimpse(df$TotalCharges)
```

> Confirmed. The values are all being treated as characters. Lets change them to numeric types.

```{r}
# changing to numeric type
df$TotalCharges <- as.numeric(df$TotalCharges)
glimpse(df$TotalCharges)
```

> Done! Just like the previous column, replace all the missing values with the median value.

```{r}
# replace missing values with median
df <- df %>%
  mutate(TotalCharges = replace(TotalCharges,
                                is.na(TotalCharges),
                                median(TotalCharges, na.rm = TRUE)))
df$TotalCharges
```

> A shorter method to changing the NA values is to start by changing the column to a numeric type. This will automatically make the NA values uniform. However this isn't always the best method given other anomalies in the data.

> More Missing Values

> This time we will look into the PaymentMethod column

```{r}
# show PaymentMethod column
df %>%
  select(PaymentMethod)
```






